Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2015-01-12 response_to_reviewers.tex

\begin{quote}
\color{blue}
We changed ``splits'' to ``shifts'' throughout the text.
\end{quote}

\textbf{18. Reviewer's Comment:}
Page 7, sentence beginning “For example, in our Nymphalidae trees...”: the authors state “we found that the split...had a variation of r = 0.5234”. Should “variation” instead be “rate”?

\textbf{18. Response:}

\begin{quote}
\color{blue}
We corrected the word from ``variation'' to ``rate''.
\end{quote}

\textbf{19. Reviewer's Comment:}
Page 7, third paragraph of Discussion: Replace “This is the reason why” with
“The relatively low phylogenetic support for many nodes in the Nymphalidae tree
is likely the reason why”.

\textbf{19. Response:}

\begin{quote}
\color{blue}
We modified the sentence according to the reviewer's suggestion.
\end{quote}
\end{document}

### 2015-01-12 MS_revised.tex
comment 18
comment 19

The MEDUSA method has been used to infer changes in net diversification
rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
of using MEDUSA on a single tree, the maximum clade credibility tree,
have been used for generation of hypotheses and discussion
\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification shifts and different
rates of diversification are found for certain lineages when
phylogenetic uncertainty was taken into account by using MEDUSA on a
random sample of trees from the posterior distribution of a Bayesian
run. We found that some diversification shifts, estimated on the
Nymphalidae maximum clade credibility tree, were found with very low
probability in the random sample of 1000 trees from the posterior
distribution (Fig. 3, Table 2). We also found that, even though the
analyses estimated the same diversification shifts on two or more trees,
the estimated net diversification rates could vary widely (Fig. 2). For
example, in our Nymphalidae trees, we found that the shift for
\emph{Magneuptychia} and \emph{Caeruleuptychia} had a variation from
\texttt{r = 0.5234}, higher than the background net diversification
rate, to almost zero. This means that observed patterns and conclusions
can be completely contradictory depending on tree choice.

@@

If there is strong phylogenetic signal for increases or decreases in net
diversification rates for a node, it is expected that these shifts would
be inferred by MEDUSA in most of the posterior distribution of trees.
However, weak phylogenetic signal for some nodes can cause some clades
to be absent in some trees and MEDUSA will be unable to estimate any
diversification shift (due to a non-existent node). This is the reason
why MEDUSA estimated net diversification rate shifts with a probability
higher than 0.90 in the sample of trees for only four shifts: the genus
\emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
shifts for other lineages with much lower probability.

>>>>

The MEDUSA method has been used to infer changes in net diversification
rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
of using MEDUSA on a single tree, the maximum clade credibility tree,
have been used for generation of hypotheses and discussion
\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification shifts and different
rates of diversification are found for certain lineages when
phylogenetic uncertainty was taken into account by using MEDUSA on a
random sample of trees from the posterior distribution of a Bayesian
run. We found that some diversification shifts, estimated on the
Nymphalidae maximum clade credibility tree, were found with very low
probability in the random sample of 1000 trees from the posterior
distribution (Fig. 3, Table 2). We also found that, even though the
analyses estimated the same diversification shifts on two or more trees,
the estimated net diversification rates could vary widely (Fig. 2). For
example, in our Nymphalidae trees, we found that the shift for
\emph{Magneuptychia} and \emph{Caeruleuptychia} had a rate from
\texttt{r = 0.5234}, higher than the background net diversification
rate, to almost zero. This means that observed patterns and conclusions
can be completely contradictory depending on tree choice.

@@

If there is strong phylogenetic signal for increases or decreases in net
diversification rates for a node, it is expected that these shifts would
be inferred by MEDUSA in most of the posterior distribution of trees.
However, weak phylogenetic signal for some nodes can cause some clades
to be absent in some trees and MEDUSA will be unable to estimate any
diversification shift (due to a non-existent node). 
The relatively low phylogenetic support for many nodes in the Nymphalidae
tree is likelyl the reason
why MEDUSA estimated net diversification rate shifts with a probability
higher than 0.90 in the sample of trees for only four shifts: the genus
\emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
shifts for other lineages with much lower probability.
